---
title: "Book case study"
output: 
  html_document:
    code_folding: "hide"

---


## Methodology

  choice: whether the customer purchased the art history of Florence. 
  1- represents a purchase
  0- representa a nonpurchase
  
Gender: 0 = Female , 1 = Male
Amount_Purchased - Total money spent on BBBC books

Frequency: Total number of purchases in the chosen period 

Last_Purchase: Months since last purchase

First_Purchase: Months since first purchase

P_Child: Number of children's books purchased

P_Youth: Number of youth books purchased

P_Cook: Number of cookbooks purchased

P_DIY: Number of do-it-yourself books purchased

P_Art: Number of art books purchased


## Loading libraries


```{r echo=FALSE}
library(readxl)
library(e1071)
library(caret)
library(ROCR)
library(ggplot2)
library(tidyverse)
library(DataExplorer)
library(ggThemeAssist)
library(esquisse)
library(gridExtra)
library(trackdown)
library(corrplot)
```



## Load in data

```{r pressure, echo=FALSE}
bbtrain = read_excel("/Users/thomasfarrell/Downloads/BBBC-Train.xlsx")
bbtest = read_excel("/Users/thomasfarrell/Downloads/BBBC-Test.xlsx")
```



## Checking for any missing values

```{r}
colSums(is.na(bbtest))
```


## Remove first column is data
We remove the first variable in the data due to the fact that this column just held a placeholder number.
```{r}
bbtrain = bbtrain[-c(1)]
bbtest = bbtest[-c(1)]
```


## Convert the response variable to a factor

```{r}
bbtrain$Choice = as.factor(bbtrain$Choice)
bbtest$Choice = as.factor(bbtest$Choice)
```



## Checking correlation within X-Variables

```{r}
b1_num = dplyr::select_if(bbtrain, is.numeric)
M = cor(b1_num)
corrplot(M, method = "number")

cor(bbtrain[sapply(bbtrain, is.numeric)])
```


Highest correlations are between First_purchase and Last_purchase, which makes sense in a way. An only customer makes a purchase and walks out, that may be both a First purchase and Last purchase observation. Repeat customers will have a longer time in months for their first First purchase than their Last purchase. 




```{r}
p1 = ggplot(data = bbtrain, mapping = aes(x = Frequency, fill = Choice))+
  geom_histogram() + theme(plot.title = element_text(face = "italic"),
    panel.background = element_rect(fill = "gray90",
        colour = "antiquewhite1", linetype = "dotted"),
    plot.background = element_rect(fill = "white",
        linetype = "dashed"))



```



```{r}

p2 = ggplot(bbtrain) +
  aes(x = Last_purchase, fill = Choice) +
  geom_histogram(bins = 30L) +
  scale_fill_hue(direction = 1) +
  theme_minimal()
```

```{r}
p3 = ggplot(bbtrain) +
  aes(x = P_Child, fill = Choice) +
  geom_histogram(bins = 30L) +
  scale_fill_hue(direction = 1) +
  theme_minimal()
```



```{r}
p4 = ggplot(bbtrain) +
  aes(x = Amount_purchased, fill = Choice) +
  geom_histogram(bins = 30L) +
  scale_fill_hue(direction = 1) +
  theme_minimal()
```




```{r}
grid.arrange(p1,p2,p3,p4)
```





## Exploratory analysis 

```{r}
par(mfrow = c(3,3))
hist(bbtrain$Amount_purchased, xlab = "Amount Purchased", main = "Amount Purchased")
hist(bbtrain$Frequency, xlab = "Total Num. Purchased", main = "Total Purchased")
hist(bbtrain$Last_purchase, xlab = "Months Since Last Purchase", main = "Last Purchase")
hist(bbtrain$First_purchase, xlab = "Months Since First Purchase", main = "First Purchase")
hist(bbtrain$P_Child, xlab = "No. Children's Books Purchased", main = "Children's Books")
hist(bbtrain$P_Youth, xlab = "No. Youth Books Purchased", main = "Youth Books")
hist(bbtrain$P_Cook, xlab = "No. Cook Books Purchased", main = "Cook Books")
hist(bbtrain$P_DIY, xlab = "No. DIY Books Purchased", main = "DIY Books")
hist(bbtrain$P_Art, xlab = "No. Art Books Purchased", main = "Art Books")
```


From these charts we can see that the Amount Purchased variable is the only variable that looks normally distributed and they other variables appear to have a right skew in their results.

## Results and findings



```{r}

```

